\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-01-30 paper/main.tex
minor updates

\subsection{Choosing a cross-validation splitting scheme}

@@

@@

In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism. Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model. However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:

\todo[inline]{Relate these to the issues listed above:}
\begin{enumerate}
	\item Build split over drugs
	\item Build split over drug-target pairs
	\item Build split over proteins
\end{enumerate} 
In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION). As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results. This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}. The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias. Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive. As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes. \\

In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION). The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein. However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier. Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE). When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures. Hence, when entirely new molecules are seen, the models perform much worse

The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.

@@

\subsection{\name identifies drugs that target a protein}

We developed \name as a computational model to predict drug--target
interactions. Specifically, given a protein, \name will identify and
rank drugs that likely target this protein. \name combines two 
types of features: structural information for drugs and proteins that
can be used to determine if they may physically interact, and

This scenario is motivated by the need to identify drugs that can target a
specific protein that may be associated with a disease or abnormal
phenotypes.
Our aim is further to combine two different types of features:
structural information based on protein structure and molecular drug composition, and their respective observable characteristics.

>>>>

\subsection{\name: computational model to identify drugs that target a
  protein}

@@

We developed \name as a computational model to predict drug--target
interactions. Specifically, given a protein, \name will identify and
rank drugs that likely target this protein. \name combines two 
types of features: structural information for drugs and proteins that
can be used to determine if they may physically interact, and
information about drug effects that may localize on an interaction
network.

@@

% In general, drug-target interaction prediction is the task of accurately predicting, whether for a given drug and a given protein there is a biological interaction within the target organism. Hereby, different training and prediction schemes lead to divergent expressiveness of the resulting model. However, when building the train-test split over compound-protein pairs for building the actual model, there are the following three options:

% \todo[inline]{Relate these to the issues listed above:}
% \begin{enumerate}
% 	\item Build split over drugs
% 	\item Build split over drug-target pairs
% 	\item Build split over proteins
% \end{enumerate} 
In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}, CITATION). As there are hopefully many more drugs to discover, the drug split scheme both emphasizes the drug repurposing idea, by applying unseen compounds to existing targets, but also benefits from more complicated drug representations, leading to tremendous\todo{???} results. This performance gain is based on minor variations among large groups of pharmaceuticals, that are easy to acquire\todo{evidence}. The second scheme has knowledge on all drugs and all proteins, and is thus prone to overfitting and the same development bias. Eventually, as there only limited drug-targets \citep{Overington2006}, predicting per protein is rather counter-intuitive. As it is hard to generalize over proteins representations, we aim at reaching similar performances for both drug and protein splitting schemes. \\

In general, recent works do perform their split over the drugs or drug-target pairs (\cite{Survey2018}% , CITATION). The first is more relevant for novel drugs, as it is much more likely to test a new compound than a innovative protein. However, it lies in the very nature of the used datasets, making the prediction for new drugs much easier. Thus, drugs are often built by minor variations of existing drugs, thus leading to no deviations in the functional group of that very compound (CITATION/EXAMPLE). When distributed over both train and test split, the models do not perform inductive inference and generalize, but rather implement transductive inference by just predicting the recently seen structures. Hence, when entirely new molecules are seen, the models perform much worse

% The same applies to splits of drug-target pairs, as all drugs were already seen, and novelty cannot be coped with.

@@
